---
title: "Computer Lab 1"
author: "Martynas Lukosevicius, Alejo Perez Gomez, Zahra Jalil Pour"
date: "03/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### 1 
It is not possible to represent exact 1/3 and 1/12 in binary. As a result it is rounded towards nearest computer float and in R its equal to 0.33333333333333331 and 0.083333333333333329 respectively.
In the first snippet code, the result is "Subtraction is wrong", but the second snippet code the result is "Subtraction is correct". 
The only numbers that represent exactly in r , are integers and fractions by power of 2 denominator. Hence all other numbers are rounded to 53 binary digits accuracy.When ever floating point operations are done , we should assume that there will be  numeric error. 1/3 and 1/12 are repeating decimals that are rounded in R.

### 2

Instead of writing  ``if(x1 - x2 == 1/12)`` it should be written ``if(isTRUE(all.equal(x1-x2,1/12)))``. In this case this equation will return TRUE.
We can use ``all.equal`` function , or we can use ``all.equal.numeric`` function too.
```{r echo = FALSE, warning=FALSE, include=FALSE}
#1.2
# all.equal.numeric() and isTRUE() function
x1 <- 1/3
x2 <- 1/4
if (isTRUE(all.equal.numeric(x1-x2, 1/12))) {
  print ("Subtraction is correct" )
} else {
  print ("Subtraction is wrong")
}
```



