---
title: "Task1"
author: "Martynas Lukosevicius, Alejo Perez Gomez, Shwetha Vandagadde Chandramouly"
date: "04/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1: Hypothesis testing

## 1.

```{r, echo = FALSE, fig.height=3}
new_df <- read.csv("lottery.csv", sep = ";")
##View(new_df)
## 1
X <- new_df$Day_of_year
Y <- new_df$Draft_No 
plot(X,Y,main="Scatterplot Y versus X", xlab="day of year ", ylab="Draft_no ", pch=19)
```

From the plot lottery looks random

## 2. 

```{r, echo = FALSE, fig.height=3}
lw1 <- loess(Y ~ X, data = new_df)
new_df$Y_hat <- predict(lw1, new_df$Day_of_year)
plot(Y ~ X, data=new_df,pch=19,cex=0.1)
lines(X,lw1$fitted,col="red",lwd=3)
```

From estimates we can see that there is some pattern in the data, however its hardly visible.

## 3.

```{r, echo = FALSE}
statis <- function(data, model){
  index_max <- which.max(model$fitted)
  index_min <- which.min(model$fitted)
  
  X_a <- data$Day_of_year[index_min]
  Y_hat_a <- model$fitted[index_min]
  
  X_b <- data$Day_of_year[index_max]
  Y_hat_b <- model$fitted[index_max]
  statistic <- (Y_hat_b - Y_hat_a)/(X_b - X_a)
  return(statistic)
}

library(boot)
fc<- function(data, index){
  data <- as.data.frame(data[index, ])
  loes <- loess(Draft_No ~ Day_of_year, data)
  return(statis(data,loes))
}


# bootstrap
results <- boot(new_df, statistic = fc, R=2000)  # i think I made mistake here about statistics
plot(results)

# h0:0, ha != 0
#data before bootstrap is not centered so we calculate distance from h0 to mean of distribution
diff <- abs(0-mean(results$t))
p_value <- (sum(results$t > (mean(results$t) + diff)) + sum(results$t < (mean(results$t) - diff)))/2000

```
$H_0: t = 0$ - lotery is random

$H_a: t \neq 0$ -lotery is not random

two sided p-value: `r p_value`

p- value is less than 0.05 so we reject null hypothesis, meaning that lotery is not random

## 4.

permutation test function:

$H_0: t = 0$ - lotery is random

$H_a: t \neq 0$ -lotery is not random

```{r}
permutation_test <- function(data, B){
  
  origin_loes <- loess(Draft_No ~ Day_of_year, data)
  t_origin <- statis(data, origin_loes)
  
  stat= numeric(B)
  n = dim(data)[1]
  for(b in 1:B){
    perm_data <- data.frame(data)
    perm_data$Day_of_year = sample(data$Day_of_year, n)
    loes_h1 <- loess(Draft_No ~ Day_of_year, perm_data)
    stat[b] <- statis(perm_data,loes_h1)
  }
  # statistic from original dat

  p_value <- sum(abs(stat) >= abs(t_origin))/B
  return(p_value)
}
```

```{r, echo = FALSE}
res <- permutation_test(new_df,2000)
```

permutation test when B = 2000: p-value = `r res`

## 5.

```{r,echo=FALSE}

new_y <- function(x, alpha){
  set.seed(12345)
  beta <- rnorm(1, 183, 10) 
  y <- max(0, min(alpha * x + beta, 366) )
  return(y)
}

#y <- sapply(new_df$Day_of_year, new_y, alpha = 0.2)

### 5-2
# new_dataset <- data.frame(Day_of_year = new_df$Day_of_year, Draft_No = y )
# permutation_test(new_dataset, 200)

### 5-3
new_alpha <-seq(0.1, 10, 0.1)
p_val <- c()
for( i in 1:length(new_alpha)){
  y <- sapply(new_df$Day_of_year,new_y, alpha = new_alpha[i])
  new_dataset <- data.frame(Day_of_year = new_df$Day_of_year, Draft_No = y )
  p_val[i]<-permutation_test(new_dataset,200)
}


```

table bellows shows p values for alpha 0.1:10 by 0.1 for which p is not equal to 0
```{r, echo=FALSE}
results <- data.frame(alpha=new_alpha,p = p_val)
dontprinttwice <- ifelse(sum(results$p !=0) == 0, print("all p values are 0"), knitr::kable(t(results[results$p !=0 , ])) )
```

power of the test is: 1-type 2 error.
type 2 error is a probability of failing to reject $H_0$ when $H_a$ is true. We know that our generated data samples are not random. The amount of rejected $H_0:$ `r sum(results$p > 0.05)`. As a result type 2 error is: `r sum(results$p > 0.05)/length(results$p) `, and power of the test is: `r 1-(sum(results$p > 0.05)/length(results$p))`.

# Question 2: Bootstrap, jackknife and confidence intervals

## 1.

```{r, echo=FALSE}
library("boot")


## 1 Plot the histogram of Price. Does it remind any conventional distribution? Compute the mean price.
prices = read.csv("prices1.csv", sep = ";")  

hist(prices$Price, breaks = 15)
```

Reminds Gamma with shape = 3, scale = 2.

## 2.

```{r,echo=FALSE}
## Mean 
mean_price <- mean(prices$Price)

## 2 Estimate the distribution of the mean price of the house using bootstrap


mean_stat <- function(data, vn){
  data<- as.data.frame(data[vn, ])
  data_mean <- as.numeric(colMeans(data['Price']))
  data_mean
}

ans <- boot(data = prices, statistic = mean_stat, R=2000)

plot(ans)

# CI percentil
ciperc <- boot.ci(ans, index = 1, type=c('perc'))

# CI bca

cibca <- boot.ci(ans, index =1, type=c('bca'))

# CI normal

cinorm <- boot.ci(ans, index =1, type=c('norm'))

## Bias-correction

T_bias_correction <- 2*mean_price - mean(ans$t)

## variance 
boot_var <- (1/(ans$R-1))* sum((ans$t-mean(ans$t))^2)

conf_int <- rbind(c(ciperc$percent[4:5]), c(cibca$bca[4:5]), c(cinorm$normal[2:3]))
row.names(conf_int) <- c("percentile", "BCa", "first-order normal")
colnames(conf_int) <- c("low", "high")
```

Bootstrap bias-correction: `r T_bias_correction`. Variance - `r boot_var`. 

95% confidence intervals: `r knitr::kable(conf_int)`


## 3.

```{r, echo=FALSE}
#3 jackknife

jackknife <- function(data, R, func){
  R <- ifelse(length(data) <= R, length(data), R)
  t <- vector()
  for (i in 1:R) {
    t[i] <- func(data[-i])
  }
  
  Ti <- R*func(data) - (R-1)*t
  var <- (1/(R*(R-1)))*sum((Ti-mean(t))^2)
  results <- list(R=R, t=t, t_mean = mean(t), variance = var)
  return(results)
}

res <- jackknife(prices$Price, 2000, mean)
```

variance using jackknife: `r res$variance` , difference between jacknife and bootstrap: `r res$variance - boot_var`. 

## 4. 

```{r, echo = FALSE}
location <- function(high,low, mean){
  length <- high-low
  loc <- (mean-low) /length
  res <- list(length = length, location = loc)
  return(res)
}

loc_ciperc <- location(ciperc$percent[5],ciperc$percent[4], mean_price)
loc_ciBCA <- location(cibca$bca[5],cibca$bca[4], mean_price)
loc_cinorm <- location(cinorm$normal[3],cinorm$normal[2], mean_price)

conf_int_comp <- rbind(c(ciperc$percent[4:5], loc_ciperc$length, loc_ciperc$location),
                  c(cibca$bca[4:5],loc_ciBCA$length, loc_ciBCA$location),
                  c(cinorm$normal[2:3], loc_cinorm$length, loc_cinorm$location))
row.names(conf_int_comp) <- c("percentile", "BCa", "first-order normal")
colnames(conf_int_comp) <- c("low", "high", "length", "location of mean")
```

Table below compares confidence intervals: `r knitr::kable(conf_int_comp)`

location of mean shows portion of interval length from the beginning of the interval to the mean  
