---
title: "Computer Lab 3"
author: "Martynas Lukosevicius"
date: "17/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Question 1

```{r }


distrib1 <- function(x,c){
  if(x>0){
    return(c*(sqrt(2*pi)^(-1))*(exp(-(c^2)/(2*x))) * (x^(-3/2)))
  }
  else return(0)
}

powerlaw <- function(x,a,t){
  if(x>t){
    return(((a-1)/t) * ((x/t)^(-a)))
  }
  else return(0)
  
}

x <- seq(0,50, by=0.01)
c <- 2
y1 <- sapply(x, distrib1, c=c)
y2 <- sapply(x, powerlaw, a = 1.3,t = 1)
ymax <- max(y1,y2)
b <- min(c(y1,y2))
e <- max(c(y1,y2))
ax <- seq(b,e,by=(e-b)/200)

plot(x,y1, type="l", col = "blue", ylim = c(0,ymax))
lines(x,y2, col = "red")
legend("topright", c("target density", "powerlaw"), fill=c("blue", "red"))
# hist(y1, breaks = ax,
#      col = "red",
#      main = "Comparison of rnorm()  with our rNorm()",
#      xlab = "values",
#      xlim = range(y1,y2))
# 
# hist(y2, breaks = ax, col = "blue",  xlim =range(y1,y2), add = TRUE)

```
Lets find maximum of target density function:
$$\frac{\partial }{\partial x}\frac{c e^{-\frac{c^2}{2 x}} x^{-\frac{3}{2}}}{\sqrt{2 \pi }}=\frac{c e^{-\frac{c^2}{2 x}} \left(c^2-3 x\right)}{2 \sqrt{2 \pi } x^{7/2}}$$
$$\frac{c \frac{e^{-c^2}}{2 x} \left(c^2-3 x\right)}{2 \sqrt{2 \pi } x^{7/2}}=0$$

$$x=\frac{c^2}{3}$$
$$T_{min} = \frac {c^2} 3$$

power-law distribution cannot be used just by itself because it doesnt support range from 0 to $T_{min}$. parameters from 0 to 1 can be taken from another distribution


lets assume that $T_{min} = 1$

probability that number will be in 0-1 region is:
$$\int_0^1 c\sqrt{2{\pi}}^{-1}\mathrm{e}^{-\frac{c^2}{2x}}{x^\frac{-3}{2}} = \dfrac{\operatorname{\Gamma}\left(\frac{1}{2},\frac{c^2}{2}\right)}{\sqrt{{\pi}}}$$
lets assume that other distribution will be uniform(0,1)

```{r }
majDensity <- function(x){
  if(x>1){
    return(powerlaw(x, 1.6, 1))
  }
  else {return(dunif(x,0,1))}
}
# y3 <- sapply(x, target)
# plot(x,y3, type = "l")
# #lines(x,dnorm(x,3,1.2), col ="red")
# lines(x,y1, col ="blue")
# #lines(x,y1, col = "red")
```

## 2

```{r }
library(poweRlaw)
randomnumber <- function(prob){
  numb <- runif(1)
  if(numb<= prob){
    return(runif(1,0,1))
  }
  else{
    return(rplcon(1,1,1.5))
  }
}

CompleteDist <- function(c, rej){
  z <- TRUE
  res <- 0
  while (z == TRUE) {
    y <- randomnumber(0.5)
    u <- runif(1)
    if(u <= distrib1(y, 2) / (c*majDensity(y))){
      res <- y
      z <- FALSE    
    }
    if(rej){
    rejected <<- rejected + 1
    }
  }
  return(res)
}

rDist <- function(n,c ,rej = FALSE){
 return(replicate(n, CompleteDist(c, rej)))
}


```

```{r , echo=FALSE}

rejected <- 0
par(mfrow=c(1,3))
datarDist <- rDist(2000, 100, TRUE)
datarDist <- datarDist[datarDist < 30]
hist(datarDist, breaks = 50)

datarDist <- rDist(2000, 1, TRUE)
datarDist <- datarDist[datarDist < 30]
hist(datarDist, breaks = 50)

datarDist <- rDist(2000, 1.5, TRUE)
datarDist <- datarDist[datarDist < 30]
hist(datarDist, breaks = 50)



# b <- min(c(datarDist))
# e <- 3
# ax <- seq(b,e,by=(e-b)/200)
# 
# hist(datarDist, breaks = 5000,
#      col = "red",
#      main = "Comparison of rnorm()  with our rNorm()",
#      xlab = "values",
#      xlim=c(0,5))


# library(ggplot2)
# # Basic histogram
# df <- as.data.frame(datarDist)
# colnames(df) <- c("weight")
# ggplot(df, aes(x=weight)) + 
#   geom_histogram(binwidth=1)+
#  geom_density(alpha=.2, fill="#FF6666") + coord_cartesian(xlim =c(0, 30), ylim = c(0, 50))
# # hist(datarnorm, breaks = ax, col = "blue",  xlim = range(datarNorm,datarnorm),  ylim = c(0,250), add = TRUE)
# # legend("topright", c("rnorm", "rNorm"), fill=c("blue", "red"))
```


```{r , echo=FALSE}
RR <- (rejected / 2000) 
```


# Question 2

## 1.

$$ DE(\mu, \alpha) = \frac{\alpha}{2} e^{-\alpha |x-\mu|}$$

* $\mu$ - location parameter
* $b > 0$ - scale parameter

inverse CDF of DE:

Source - https://en.wikipedia.org/wiki/Laplace_distribution

$$F^{-1}(p) = \mu - b sgn(p-0.5)ln(1-2|p-0.5|)$$

where $b = \frac 1 \alpha$


```{r }
rLaplace <- function(n, mean = 0, alpha = 1){
  b <- 1/alpha
  u <- runif(n)
  res <- mean - (b*sign(u-0.5) * log(1-(2*abs(u-0.5))))
  return(res)
}
```

meaning:

1. calculate b.
2. take n random variables from uniform distribution [0,1]
3. calculate random numbers from inverse CDF of laplace distribution where x is a random variable from uniform distribution


```{r , echo=FALSE, warning=FALSE, message=FALSE}
library(rmutil)
c1 <- rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

hist(rlaplace(10000, 2,1), 100, 
     col = c1,
     main = "Comparison of rlaplace function from rmutil with our rLaplace", 
     xlab = "values")
hist(rLaplace(10000, 2,1), 100, col = c2, add = TRUE)
legend("topright", c("rLaplace", "rlaplace"), fill=c(c2, c1))
```


## 2.

```{r }
DE <- function(x, mean = 0,alpha = 1){
  return((0.5*alpha)*exp((-alpha)*abs(x-mean)))
}

genNorm <- function(c, rej){
  z <- TRUE
  res <- 0
  while (z == TRUE) {
    y <- rLaplace(1)
    u <- runif(1)
    if(u <= pnorm(y) / (c*DE(y))){
      res <- y
      z <- FALSE    
    }
    if(rej){
    rejected <<- rejected + 1
    }
  }
  return(res)
}
  
rNorm <- function(n,c,rej = FALSE){
 return(replicate(n, genNorm(c, rej)))
}
```

algorithm:

1. write Laplace probability function
2. assign 0 to result value res
3. generate random number y from rLaplace function
4. generate random number u from uniform distribution 
5. check if u is less or equal to probability of y in normal distribution / c * probability of y in laplace distribution
    a) if yes, return y
    b) repeat steps from 3


```{r , echo=FALSE}
# x <- seq(-5,5, by=0.1)
# 
# test <- function(x){ return((dnorm(x)/DE(x)))}
# cde <- sapply(x, test)
# cnorm <- dnorm(x)
#c <- max(cnorm/cde)

## why it calculates 1,35 ???
```

$$c>0;sup_x(f(x)/g(x)) \le c$$

$$h(x) = \frac {f(x)}  {g(x)}$$
$$f(x) = N(0,1) = \frac 1 {\sqrt{2 \pi}} e^{ - \frac {x^2} 2}$$
$$g(x) = \frac 1 2 e^{-|x|}$$
$$h(x) = \sqrt{\frac{2}{\pi }} e^{| x| -\frac{x^2}{2}}$$

$$\frac{d}{x d}\text{Null}\sqrt{\frac{2}{\pi }} e^{| x| -\frac{x^2}{2}}=\sqrt{\frac{2}{\pi }} x e^{| x| -\frac{x^2}{2}} \left(\frac{1}{| x| }-1\right)$$

$$\frac{\sqrt{2} e^{x-\frac{x^2}{2}} (x-1)}{\pi }=0$$
$$x=\pm1$$
$$c = h(1) = 1.3154892$$


source  - http://www.columbia.edu/~ks20/4703-Sigman/4703-07-Notes-ARM.pdf


```{r , echo=FALSE}

c <- 1.3154892

rejected <- 0
datarNorm <- rNorm(2000, c, TRUE)
datarnorm <- rnorm(2000)

b <- min(c(datarNorm,datarnorm))
e <- max(c(datarNorm,datarnorm))
ax <- seq(b,e,by=(e-b)/50)



hist(datarNorm, breaks = ax,
     col = c1,
     main = "Comparison of rnorm()  with our rNorm()",
     xlab = "values",
     xlim = range(datarNorm,datarnorm))

hist(datarnorm, breaks = ax, col = c2,  xlim = range(datarNorm,datarnorm),  ylim = c(0,250), add = TRUE)
legend("topright", c("rnorm", "rNorm"), fill=c(c2, c1))
```

```{r , echo=FALSE}
compare <- matrix(c(mean(datarNorm), var(datarNorm), mean(datarnorm), var(datarnorm)), nrow = 2, byrow = TRUE)
row.names(compare) <- c("rNorm()", "rnorm()")
colnames(compare) <- c("mean", "variance")
knitr::kable(compare)
```

```{r , echo=FALSE}
RR <- (rejected / 2000) 
```

rejection rate: `r 1-(1/RR)`, expected rejection rate = c = `r 1-(1/c)`, difference - `r RR - c`
