---
title: "Computer Lab 3"
author: "Martynas Lukosevicius, Alejo Perez Gomez, Zahra Jalil Pour"
date: "17/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Question 1

```{r , echo=FALSE}


distrib1 <- function(x,c){
  if(x>0){
    return(c*(sqrt(2*pi)^(-1))*(exp(-(c^2)/(2*x))) * (x^(-3/2)))
  }
  else return(0)
}

powerlaw <- function(x,a,t){
  if(x>t){
    return(((a-1)/t) * ((x/t)^(-a)))
  }
  else return(0)
  
}

x <- seq(0,50, by=0.01)
c <- 1.6
y1 <- sapply(x, distrib1, c=c)
y2 <- sapply(x, powerlaw, a = 2,t = 3)
ymax <- max(y1,y2)
b <- min(c(y1,y2))
e <- max(c(y1,y2))
ax <- seq(b,e,by=(e-b)/200)

plot(x,y1, type="l", col = "blue", ylim = c(0,ymax))
lines(x,y2, col = "red")
legend("topright", c("target density", "powerlaw"), fill=c("blue", "red"))


```
Power-law distribution cannot be used just by itself because it does not support range from 0 to $T_{min}$. Because of this, we need to use mixture distribution. To support x from 0 to $T_{min}$ we choose uniform distribution $Unif(0,T_{min})$. As we can see, powerlaw distribution is monotonically decreasing, so $T_{min}$ should be equal to x where target density has maximum value.

Let's find maximum of target density function:

$$\frac{\partial }{\partial x}\frac{c e^{-\frac{c^2}{2 x}} x^{-\frac{3}{2}}}{\sqrt{2 \pi }}=\frac{c e^{-\frac{c^2}{2 x}} \left(c^2-3 x\right)}{2 \sqrt{2 \pi } x^{7/2}}$$
$$\frac{c \frac{e^{-c^2}}{2 x} \left(c^2-3 x\right)}{2 \sqrt{2 \pi } x^{7/2}}=0$$

$$x=\frac{c^2}{3}$$
$$T_{min} = \frac {c^2} 3$$




<!-- lets assume that $T_{min} = 1$ -->

To make a mixture model we need to know the probability of taking uniform distribution and powerlaw distribution probability. That number will be in the range $[0,T_{min}]$:

$$\int_0^{T_{min}} c x^{-\frac{3}{2}} e^{-\frac{c^2}{2 x}} {\sqrt{2 \pi }}^{-1}  dx = \dfrac{\operatorname{\Gamma}\left(\frac{1}{2},\frac{c^2}{2T_{min}}\right)}{\sqrt{{\pi}}}$$

as $T_{min} = \frac {c^2} 3$ - $\dfrac{\operatorname{\Gamma}\left(\frac{1}{2},\frac{c^2}{2T_{min}}\right)}{\sqrt{{\pi}}} = \frac{\Gamma \left(\frac{1}{2},\frac{3}{2}\right)}{\sqrt{\pi }} \sim 0.08326451666$


As a result, majorising density function is:

$$
g(x) =  \frac{2 * 0.08326}{c^2} 1_{[0,T_{min}]} + (1-0.08326)* \frac{2^{1-a} (a-1) \left(\frac{x}{c^2}\right)^{-a}}{c^2} * 1_{(T_{min},\infty)}
$$


## 2. 

Target density:

$$f(x) = c(\sqrt{2\pi})^{-1} e^{\frac {-c^2} {2x}} x^{ \frac {-3}{2}} 1_{(0,\infty)}(x)$$

We need to find $c_{maj}$

$$c_{maj}>0;sup_x(f(x)/g(x)) \le c_{maj}$$

$$h(x) = \frac {f(x)}  {g(x)}$$
$c_{maj} = h(x_{maj})$

if x < $T_{min}$ 

$$\frac {\partial}{\partial x} \frac{f(x)}{\frac{0.16652}{c^2}} = \frac{e^{-\frac{c^2}{2 x}} \left(1.19788 c^5-3.59364 c^3 x\right)}{x^{7/2}}$$

$$\frac{e^{-\frac{c^2}{2 x}} \left(1.19788 c^5-3.59364 c^3 x\right)}{x^{7/2}} = 0$$

$$x_{maj}=\frac{c^2}{3}$$

x for cmaj = $0.333333 c^2 = T_{min}$



if x> Tmin

$$\frac {\partial}{\partial x} \frac{f(x)}{g(x)} = \frac{3^{a-1} c^3 \frac{e^{-c^2}}{2 x} \left(\frac{x}{c^2}\right)^a \left((2 a-3) x+c^2\right)}{2 \sqrt{2 \pi } (a-1) x^{7/2}}$$
$$\frac{3^{a-1} c^3 \frac{e^{-c^2}}{2 x} \left(\frac{x}{c^2}\right)^a \left((2 a-3) x+c^2\right)}{2 \sqrt{2 \pi } (a-1) x^{7/2}} = 0$$

$$ x_{maj}=\frac{c^2}{3-2 a}$$
when $1<a<1.5$

$$c_{maj} = h(x_{maj})$$


As we can see $a$ should be between 1 and 1.5 If $a >=1.5$ majorising constant will be negative. We will fix $\alpha=1.3$. Afterwards, the majorizing constant will ensure it gets above of the target.

```{r }
majDensity <- function(x, c, a){
  Tmin <- (c^2)/3
  
  if(x>Tmin){
    return((1-0.083264) * powerlaw(x, a, Tmin))
  }
  else {return(0.083264 * dunif(x,0,Tmin))}
}
c <- 2
y3 <- sapply(x, majDensity, c = c, a = 1.3)
y1 <- sapply(x, distrib1, c = c)
plot(x,y1, type = "l", ylim = c(0,max(y3,y1)), ylab = "y")
lines(x,y3, col ="blue")
legend("topright", c("target", "proposal, a = 1.3"), fill=c("black", "blue"))

```

## 2

```{r , message=FALSE, warning=FALSE, error=FALSE}
library(poweRlaw)
randomnumber <- function(t,a){
  numb <- runif(1)
  if(numb<= 0.08326){
    return(runif(1,0,t))
  }
  else{
    return(rplcon(1,t,a))
  }
}

CompleteDist <- function(c, a, rej){
  z <- TRUE
  res <- 0

  Tmin <- (c^2)/3
  xmaj <- (c^2)/(3-2*a)
  cmaj <- distrib1(xmaj,c)/majDensity(xmaj,c, a) 
  
  while (z == TRUE) {
    y <- randomnumber(Tmin,a)
    u <- runif(1)
    if(u <= distrib1(y, c) / (cmaj*majDensity(y,c,a))){
      res <- y
      z <- FALSE    
    }
    if(rej){
    rejected <<- rejected + 1
    }
  }
  return(res)
}

rDist <- function(n,c,a ,rej = FALSE){
 return(replicate(n, CompleteDist(c, a, rej)))
}


```

## 3.

```{r , echo=FALSE}


rejected <- 0
par(mfrow=c(1,3))
datarDist1 <- rDist(2000, 2, 1.3, TRUE)
datarDist12 <- datarDist1[datarDist1 < 200]
rejected1 <- rejected

rejected <- 0
datarDist2 <- rDist(2000, 9, 1.3, TRUE)
datarDist22 <- datarDist2[datarDist2 < 200]
rejected2 <- rejected

rejected <- 0
datarDist3 <- rDist(2000, 15, 1.3, TRUE)
datarDist32 <- datarDist3[datarDist3 < 200]
rejected3 <- rejected

hist(datarDist12, breaks = 30, main = "c = 2", xlab = "data")
hist(datarDist22, breaks = 30, main = "c = 9", xlab = "data")
hist(datarDist32, breaks = 30, main = "c = 15", xlab = "data")



results <- matrix(c(mean(datarDist1), var(datarDist1), mean(datarDist2), var(datarDist2), mean(datarDist3), var(datarDist3)), nrow = 2)
colnames(results) <- c("c=2", "c=9", "c=15")
row.names(results) <- c("mean","variance")

```
```{r , echo=FALSE}
knitr::kable(results) 
```

Variance and mean is increase due to c

```{r , echo=FALSE}

resrejected <- numeric()
c <- seq(1.1,20, by = 0.1)
for (ci in c) {
  rejected <- 0
  rDist(2000, ci, 1.12, TRUE)
  RR <- (rejected / 2000)
  resrejected <- append(resrejected ,1-(1/RR))
}

```

```{r , echo=FALSE}
plot(c,resrejected, type="l", ylab= "rejection rate")
```

From plot we can see that rejection rate does not depend on the constant c of the target density. 

# Question 2

## 1.

$$ DE(\mu, \alpha) = \frac{\alpha}{2} e^{-\alpha |x-\mu|}$$

* $\mu$ - location parameter
* $b > 0$ - scale parameter

inverse CDF of DE:

Source - https://en.wikipedia.org/wiki/Laplace_distribution

$$F^{-1}(p) = \mu - b sgn(p-0.5)ln(1-2|p-0.5|)$$

where $b = \frac 1 \alpha$


```{r }
rLaplace <- function(n, mean = 0, alpha = 1){
  b <- 1/alpha
  u <- runif(n)
  res <- mean - (b*sign(u-0.5) * log(1-(2*abs(u-0.5))))
  return(res)
}
```

meaning:

1. calculate b.
2. take n random variables from uniform distribution [0,1]
3. calculate random numbers from inverse CDF of laplace distribution where x is a random variable from uniform distribution


```{r , echo=FALSE, warning=FALSE, message=FALSE}
library(rmutil)
c1 <- rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

hist(rlaplace(10000, 2,1), 100, 
     col = c1,
     main = "Comparison of rlaplace function from rmutil with our rLaplace", 
     xlab = "values")
hist(rLaplace(10000, 2,1), 100, col = c2, add = TRUE)
legend("topright", c("rLaplace", "rlaplace"), fill=c(c2, c1))
```
From histogram we can see that inverse CDF method approximated Laplace distribution reasonably, mean and variance similarly, in turn.

## 2.

```{r }
DE <- function(x, mean = 0,alpha = 1){
  return((0.5*alpha)*exp((-alpha)*abs(x-mean)))
}

genNorm <- function(c, rej){
  z <- TRUE
  res <- 0
  while (z == TRUE) {
    y <- rLaplace(1)
    u <- runif(1)
    if(u <= pnorm(y) / (c*DE(y))){
      res <- y
      z <- FALSE    
    }
    if(rej){
    rejected <<- rejected + 1
    }
  }
  return(res)
}
  
rNorm <- function(n,c,rej = FALSE){
 return(replicate(n, genNorm(c, rej)))
}
```

algorithm:

1. write Laplace probability function
2. assign 0 to result value res
3. generate random number y from rLaplace function
4. generate random number u from uniform distribution 
5. check if u is less or equal to probability of y in normal distribution / c * probability of y in laplace distribution
    a) if yes, return y
    b) repeat steps from 3


```{r , echo=FALSE}
# x <- seq(-5,5, by=0.1)
# 
# test <- function(x){ return((dnorm(x)/DE(x)))}
# cde <- sapply(x, test)
# cnorm <- dnorm(x)
#c <- max(cnorm/cde)

## why it calculates 1,35 ???
```

$$c>0;sup_x(f(x)/g(x)) \le c$$

$$h(x) = \frac {f(x)}  {g(x)}$$
$$f(x) = N(0,1) = \frac 1 {\sqrt{2 \pi}} e^{ - \frac {x^2} 2}$$
$$g(x) = \frac 1 2 e^{-|x|}$$
$$h(x) = \sqrt{\frac{2}{\pi }} e^{| x| -\frac{x^2}{2}}$$

$$\frac{d}{dx }\sqrt{\frac{2}{\pi }} e^{| x| -\frac{x^2}{2}}=\sqrt{\frac{2}{\pi }} x e^{| x| -\frac{x^2}{2}} \left(\frac{1}{| x| }-1\right)$$

$$\frac{\sqrt{2} e^{x-\frac{x^2}{2}} (x-1)}{\pi }=0$$
$$x=\pm1$$
$$c = h(1) = 1.3154892$$


```{r , echo=FALSE}

c <- 1.3154892

rejected <- 0
datarNorm <- rNorm(2000, c, TRUE)
datarnorm <- rnorm(2000)

b <- min(c(datarNorm,datarnorm))
e <- max(c(datarNorm,datarnorm))
ax <- seq(b,e,by=(e-b)/50)



hist(datarNorm, breaks = ax,
     col = c1,
     main = "Comparison of rnorm()  with our rNorm()",
     xlab = "values",
     xlim = range(datarNorm,datarnorm))

hist(datarnorm, breaks = ax, col = c2,  xlim = range(datarNorm,datarnorm),  ylim = c(0,250), add = TRUE)
legend("topright", c("rnorm", "rNorm"), fill=c(c2, c1))
```

```{r , echo=FALSE}
compare <- matrix(c(mean(datarNorm), var(datarNorm), mean(datarnorm), var(datarnorm)), nrow = 2, byrow = TRUE)
row.names(compare) <- c("rNorm()", "rnorm()")
colnames(compare) <- c("mean", "variance")
knitr::kable(compare)
```


```{r , echo=FALSE}
RR <- (rejected / 2000) 
```

rejection rate: `r 1-(1/RR)`, expected rejection rate = $1 - \frac 1 c$ = `r 1-(1/c)`, difference - `r (1-(1/RR)) - (1-(1/c))`

We can see that mean and variance of generated distribution slightly differs.

# Appendix

```{r ref.label=knitr::all_labels(), echo = T, eval = F}
```
